Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-13 thesis/chapters/ch_introduction.tex
star formation, star particle section updates. radiation updates. moving hydro into new section. preface to ingredients of GE section

Very broadly give a picture of galactic evolution, focusing in on the physics needed to simulate galaxies (cooling, heating, chemistry, feedback). Touch on inflows, outflows, mixing, ISM physics, etc. Bring up CE and feedback obviously but reserve for next section(s). But maybe this can all go in the intro-intro

\subsection{Nucleosynthesis and Stellar Evolution}

@@

\subsection{Star Formation and Stellar Feedback}\label{sec:section1}

As elements are synthesized throughout stellar evolution, any framework for modelling galactic evolution requires an understanding of both where, how, and with what properties stars form in a galaxy.

\subsubsection{Star Formation}

In galaxy-scale hydrodynamics, star formation is typically treated simplistically.

@@

\subsubsection{Stars in Simulations}

@@

\subsubsection{Stellar Radiation Feedback}
\ref{sec:radiation}

We only briefly discuss this method of feedback here, as it is discussed in much more detail in both Chapters~\ref{ch:chapter_1} and \ref{ch:chapter_2}. In brief, \cite{Agertz2013} demonstrated that stellar radiation, accounts for a substantial fraction of the energy output of a stellar population. Ionizing photons contribute

\section{Semi-Analytic Models / One-zone Models}

Definitely have a lot here. Motivate need for better understanding of mixing processes in galaxies. \citep{SchonrichWeinberg2019}

\section{Hydrodynamics}
\textit{(This will be a sub-section of something, but putting here for now as placeholder).}

In this Dissertation, we make use of the adaptive-mesh refinement (AMR) hydrodynamics code \code{Enzo}, as described in greater detail in Chapter~\ref{ch:chapter_1}. As it pertains to galactic chemical evolution, however, the use of a grid-based code has the advantage that metals (which are advected as passive scalars that follow the fluid-flow in the simulations) are allowed to mix and diffuse naturally across fluid elements (grid cells). This is not possible in native SPH implementations, yet is necessary to reproduce realistic galactic chemical evolution properties as shown by multiple recent works testing implementations of diffusion in SPH simulations \citep[e.g.][]{Shen2010,Su2017a,Escala2018}. The disadvantage of the diffusion in \code{Enzo}, however, is that it is entirely numerical, and therefore its exact properties are challenging to characterize and are strongly resolution dependent. Throughout this dissertation we attempt to account for the effects of resolution on our results, but do not explicitly examine the properties of numerical diffusion itself.

\section{Structure of Dissertation}\label{sec:structure}

@@

In Chapter \ref{ch:chapter2} \citep[published as ][]{Emerick2018a}, we investigate how stellar ionizing radiation feedback impacts the evolution of our simulated low mass dwarf galaxy. We confirm the findings in recent works, as dicsussed in Section~\ref{sec:radiation}, that find that stellar radiation is an important source of pre-supernova feedback that regulates star formation and can help drive galactic outflows outflows. We additionally investigate how radiation feedback couples to the ISM to drive galactic winds by demonstrating that the ionization of gas far from individual ionzing sources, out into the disk-halo interface of our galaxy, creates low-density channels through which supernovae can readily drive gas and metals into and beyond the galaxy's CGM. This suggests that not only is the local deposition of energy from stellar radiation important, which quickly destroys dense gas in stellar birth-sites, but also long-range effects are important to consider in a consistent model of stellar feedback.

>>>>

We give a primer on how to model galactic evolution in hydrodynamics simulations. We focus on the physics needed to simulate galaxies (cooling, heating, chemistry, star formation, feedback, etc.) as it pertains to the aspects of galactic chemical evolution examined in this dissertation. This is meant to be a broad overview of the physical processes modelled in such simulations, focusing on aspects of their implementation rather than the derivation of theory behind such processes. While important, the latter is better reserved for a much more detailed work (i.e. textbook), and focusing on the former gives a much more accurate represenation of the actual work done for this dissertation. This discussion includes a very brief overview of hydrodynamics and the numerical methods used in this work in Section~\ref{intro:sec:hydrodynamics}, nucleosynthesis and stellar evolution in Section~\ref{intro:sec:nucleosynthesis}, star formation and star particles in Section~\ref{intro:sec:sf} and Section~\ref{intro:sec:stars},

\subsection{Hydrodynamics}
\textit{(This will be a sub-section of something, but putting here for now as placeholder).}

In this Dissertation, we make use of the adaptive-mesh refinement (AMR) hydrodynamics code \code{Enzo}, as described in greater detail in Chapter~\ref{ch:chapter1}. As it pertains to galactic chemical evolution, however, the use of a grid-based code has the advantage that metals (which are advected as passive scalars that follow the fluid-flow in the simulations) are allowed to mix and diffuse naturally across fluid elements (grid cells). This is not possible in native SPH implementations, yet is necessary to reproduce realistic galactic chemical evolution properties as shown by multiple recent works testing implementations of diffusion in SPH simulations \citep[e.g.][]{Shen2010,Su2017a,Escala2018}. The disadvantage of the diffusion in \code{Enzo}, however, is that it is entirely numerical, and therefore its exact properties are challenging to characterize and are strongly resolution dependent. Throughout this dissertation we attempt to account for the effects of resolution on our results, but do not explicitly examine the properties of numerical diffusion itself.

\subsection{Nucleosynthesis and Stellar Evolution} \label{intro:sec:nucleosynthesis}

@@

%\subsection{Star Formation and Stellar Feedback}\label{intro:sec:sffb}
%
%As elements are synthesized throughout stellar evolution, any framework for modelling galactic evolution requires an understanding of both where, how, and with what properties stars form in a galaxy.

\subsection{Star Formation} \label{intro:sec:sf}

In galaxy-scale hydrodynamics, star formation is typically treated as a process that occurs by a sub-grid model, with an assumed rate and assumed efficiency, in fluid elements that surpass a variety of threshold conditions. Usually, simulators set these thresholds to fully encompass gas whose Jeans-length (and thus dynamical behavior) becomes unresolved according to the widely-used \cite{Truelove1997} criterion. Often, this means a simple mass or number density threshold, on top of which simulators can also require that the local gas cells be in a converging flow ($\div \bm{v} < 0$), that gas surpass some virial threshold \citep[e.g.][]{}, or that gas also meet a minimum $H_2$ mass fraction \citep[e.g.][]{}. One a fluid element is flagged as star forming, simulators have a few options to choose from in how to determine what fraction and with what rate that mass is converted into stars. Briefly, some form of the following equation is usually adopted,
\begin{equation}
\label{intro:eq:sfr}
 \rm{SFR} \propto 
\end{equation}
,. Some works adopt modifications of this simple perscription, with varying star formation efficiency ($\eta$) (refs), or XXXX. Others, for application in high-resolution simulations where the mass of stars that would form in a single timestep ($dt \times \rm{SFR}$) is far less than a single star particle mass, allow stars to form out of star forming fluid elements stochastically \citep[e.g.][]{Goldbaum2015}.

Recent work has demonstrated that variations in these implemenations can lead to dramatic differences in the outcome of the simulated galaxies (refs, incl brooks/munshi work). However, it is generally found that the differences between implementations decrease in simulations with sufficient resolution ($\sim$ 10~pc or better spatial, or $\sim$ $100$ M$_{\odot}$ or better in mass) and with a sufficiently high density threshold for star formation ($n > 100$~cm$^{-3}$) (fire ref).

@@

\subsection{Stars in Simulations} \label{intro:sec:stars}

@@

\subsection{Stellar Feedback} \label{intro:sec:feedback}

\subsubsection{Stellar Radiation Feedback}
\label{intro:sec:radiation}

We only briefly discuss this method of feedback here, as it is discussed in much more detail in both Chapters~\ref{ch:chapter1} and \ref{ch:chapter2}. In brief, \cite{Leitherer1999}, \cite{Agertz2013}, and others demonstrated that stellar radiation, accounts for a substantial fraction of the total energy output of a stellar population. Ionizing photons from young, massive stars can photoionize and photoheat the gas surrounding sites of recent star formation. This effectively pre-processes the local environment within which SNe occur, often increasing the effectiveness of SNe in regulating star formation and driving outflows \citep{Hu2016}. In addition, radiation pressure in the single-scattering limit (from UV photons) and rescattered infrared photons has been used as an additional source of pre-SN feedback \citep[e.g.][]{FIRE}. However, due in part to differences in how ionizing radition is followed across simulations, there is disagreement as to exactly how and in what conditions this acts as a source of feedback (see \cite{Krumholz2018} for a detailed examination of this problem).

Non-ionizing radiation also contributes to the regulation of star formation and the multi-phase ISM. Far ultraviolet (FUV) photons contribute to gas heating via the photoelectric heating of dust grains. This is a significant heating mechanism in higher metallicity environments, like the Milky Way \citep{Parravano2003,Wolfire2003}, but may still be important in setting temperature floor of cold dense gas in lower mass galaxies \citep{Forbes2016,Hu2017}. Lyman-Werner (LW) band radiation regulates the $H_2$ abundance, which is particularly important in low-metallicity environments where $H_2$ is a dominant coolant \citep[e.g.][]{Wolcott-Green2012}.\footnote{And is arguably even more important when using star formation perscriptions that depend upon the local $H_2$ mass fraction.} As the goal of this dissertation is to develop an as-complete-as-possible model of galactic chemical evolution, we account for each of the processes in some fashion, including the use of an adaptive ray-tracing radiative transfer scheme to track the stellar ionizing photons.

%
%
%
%\section{Semi-Analytic Models / One-zone Models}
%
%Definitely have a lot here. Motivate need for better understanding of mixing processes in galaxies. \citep{SchonrichWeinberg2019}
%
%

\section{Structure of Dissertation}\label{intro:sec:structure}

@@

In Chapter \ref{ch:chapter2} \citep[published as ][]{Emerick2018a}, we investigate how stellar ionizing radiation feedback impacts the evolution of our simulated low mass dwarf galaxy. We confirm the findings in recent works, as dicsussed in Section~\ref{intro:sec:radiation}, that find that stellar radiation is an important source of pre-supernova feedback that regulates star formation and can help drive galactic outflows outflows. We additionally investigate how radiation feedback couples to the ISM to drive galactic winds by demonstrating that the ionization of gas far from individual ionzing sources, out into the disk-halo interface of our galaxy, creates low-density channels through which supernovae can readily drive gas and metals into and beyond the galaxy's CGM. This suggests that not only is the local deposition of energy from stellar radiation important, which quickly destroys dense gas in stellar birth-sites, but also long-range effects are important to consider in a consistent model of stellar feedback.
